### 2017-04-21 sheets/basis-1/sheet-basis-1.tex
Template for exercise sheets

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 0}{Sets, functions, groups and orders}

\section{Sets and functions}

\begin{exercise}
  The composition of injective functions is injective.
\end{exercise}

\begin{exercise}
  If $f \circ g$ is surjective, what do you know about $g$? Is it injective, surjective, bijective?
\end{exercise}

\section{Semigroups, monoids and groups}

\begin{definition}{Semigroup}
  A semigroup $(S, \cdot)$ consists out of a non-empty type $S$ with an operation $\cdot : S \to S \to S$ which is associative.
  $S$ is called the carrier.
\end{definition}

\begin{definition}{Monoid}
  A monoid is a semi-group with an identity $\alpha$, i.e. $\forall x : S, \alpha \cdot x = x = x \cdot \alpha$.
\end{definition}

\begin{exercise}
  List semi-groups with the following carriers:
  \begin{multicols}{3}
    \begin{itemize}
    \item $\mathbb{N}$
    \item $\mathcal{P}(W)$ for a finite set $W$
    \item $S$ for any non-empty set $S$
    \end{itemize}
  \end{multicols}

  Which of them is a monoid?
\end{exercise}

\begin{definition}{Canonical pre-order}
  A semi-group can be ordered in two canonical ways:
  \begin{itemize}
  \item $x \leq^R_\cdot y := \exists z, y = x \cdot z$
  \item $x \leq^L_\cdot y := \exists z, y = z \cdot x$
  \end{itemize}
\end{definition}

\begin{definition}{Canonically ordered semi-group}
  A semi-group is canonically ordered if $\leq^R_\cdot$ and $\leq^L_\cdot$ are partial orders.
\end{definition}

\begin{definition}{Group}
  A group is a monoid $(S, \cdot, \alpha)$ with inverse elements, i.e. $\forall x, \exists x^{-1}, x \cdot x^{-1} = \alpha$.
  A group is trivial if $S$ is a singleton.
\end{definition}

\begin{exercise}
  Only a trivial group is canonically ordered.
\end{exercise}

>>>>

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 1}{What is a Category?}

\begin{hint}
  Read Chapter 1.1. For initial and terminal objects refer to Definition 2.17 on pages 48ff. For Monics and Epics, see the respective sections in Chapters 5.1 and 5.2.
\end{hint}
